---
title:  "Finding LS estimates in $R$"
author: Dr. Hasthika Rupasinghe
date: "August 22th 2023"
output:
  rmdformats::readthedown:
    toc_depth: 3
---

```{r setup, include=FALSE, comment=NA, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Read data from an URL


```{r comment=NA, warning=FALSE, message=FALSE}
toluca <- read.table("http://www.cnachtsheim-text.csom.umn.edu/Kutner/Chapter%20%201%20Data%20Sets/CH01TA01.txt", sep ="" , header = FALSE)


#Look at the first 6 entries
head(toluca)
```


### Rename columns 
<!-- This is how I should rename my x and y variables in the dataset-->

```{r comment=NA, warning=FALSE, message=FALSE}
colnames(toluca) <- c("lotSize", "hours")

#Look at the first 6 entries
#head(toluca)

```

### Creating a scatter plot 
<!--  Make sure x and y variables are quantitative--> 

```{r comment=NA, warning=FALSE, message=FALSE}
library(ggplot2)
ggplot(toluca, aes(x = lotSize, y = hours)) +
  geom_point() +
  labs(x = "Lot Size", y = "Work Hours", title = "Toluca example scatter plot") +
  theme_bw()
```
Note: Lot Size and Work hours has a strong, linear, positive association



### Creating a scatter plot, LS line added

```{r comment=NA, warning=FALSE, message=FALSE}
ggplot(toluca, aes(x = lotSize, y = hours)) +
  geom_point() +
  labs(x = "Lot Size", y = "Work Hours", title = "Toluca example, LS line added") +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()
```

### Finding the LS estimates
```{r comment=NA, warning=FALSE, message=FALSE}
toluca_LS_model <- lm(hours ~ lotSize, data = toluca)
summary_toluca_LS_model <- summary(toluca_LS_model)
summary_toluca_LS_model
```

### Finding fitted values $\hat{y_i}$, and residuals $e_i =(y_i-\hat{y_i})$

```{r comment=NA, warning=FALSE, message=FALSE}
#Method 1
#library(moderndive)
#Fittedandresiduals <-get_regression_points(toluca_LS_model)
#Fittedandresiduals

#Method 2
#Simply use the summary_toluca_LS_model object to get residuals and fitted values
summary_toluca_LS_model$residuals
```

### Calculating $SSE = \sum (y_i-\hat{y_i})^2$

```{r comment=NA, warning=FALSE, message=FALSE}
#Using method 1 above
#sum_of_square_of_residuals <- sum(Fittedandresiduals$residual^2)
#sum_of_square_of_residuals

#Using method 2
sum_of_square_of_residuals <- sum(summary_toluca_LS_model$residuals^2)
sum_of_square_of_residuals
```

### Calculating $MSE =SSE/(n-2)$

```{r comment=NA, warning=FALSE, message=FALSE}
Mean_Square_Error <- sum_of_square_of_residuals/(nrow(toluca) -2)
Mean_Square_Error

```

### Calculating Residual Standard Error (estimator of standard deviation $\sigma$) $s = \sqrt(MSE)$

```{r}
s <- sqrt(Mean_Square_Error)
s
```
