---
title: "Multiple Regression in $R$"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE, comment=NA, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We’ll use the `Credit` dataframe from the `ISLR` package to demonstrate multiple regression with:

A numerical outcome variable $y$, in this case credit card balance.

Two explanatory variables:

  1) A first numerical explanatory variable $x_1$. In this case, their credit limit.

  2) A second numerical explanatory variable $x_2$. In this case, their income (in thousands of dollars).

Note: This dataset is not based on actual individuals, it is a simulated dataset used for educational purposes.

Analogous to simple linear regression, where the regression function $E\{Y\} = \beta_0 + \beta_1X_1$ is
a line, regression function $E\{Y\} = \beta_0 + \beta_1X_1 +\beta_2X_2$ is a plane.

```{r comment=NA, warning=FALSE, message=FALSE}
library(ISLR)
library(plotly)

data(Credit)
head(Credit)

# draw 3D scatterplot
p <- plot_ly(data = Credit, z = ~Balance, x = ~Income, y = ~Limit, opacity = 0.6, color = Credit$Balance) %>%
  add_markers() 
p

#Fit a multiple linear regression model 
#Syntax: model_name <- lm(y ~ x1 + x2 + ... +xn, data = data_frame_name)

Balance_model <- lm(Balance ~ Limit + Income, data = Credit)
Balance_model
range(Credit$Limit)
range(Credit$Income)
```

## Scatter plot matrix for Dwaine Studios example {-}


```{r}
#DwaineData <- read.table("https://people.stat.sc.edu/Hitchcock/studiodata.txt", sep ="" , header = FALSE)

#DwaineData <- read.table("http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%206%20Data%20Sets/CH06FI05.txt", sep ="" , header = FALSE)

DwaineData <- read.table("https://people.stat.sc.edu/Hitchcock/studiodata.txt", sep ="" , header = FALSE)

DwaineData <- data.frame(DwaineData)


colnames(DwaineData) <- c("TARGTPOP", "DISPOINC", "SALES")
head(DwaineData)
dim(DwaineData)

pairs(DwaineData)




```


## Correlation matrix for Dwaine Studios example {-}

```{r}
cor(DwaineData)
```


```{r}
# Wanna make it fancy?
library("GGally")
ggpairs(DwaineData)
```


## Breusch-Pagan Test for Constancy of Error Variance {-}

```{r comment=NA, warning=FALSE, message=FALSE}
DwaineModel <- lm(SALES ~ TARGTPOP + DISPOINC, data = DwaineData)

library(lmtest)
bptest(DwaineModel, student = FALSE)
```

### F test for lack of fit {-}

```{r}
# F test for lack of fit

#Reduced <- lm(SALES ~ TARGTPOP + DISPOINC, data = DwaineData) # This is our usual model
#summary(Reduced)



#Full <-  lm(SALES ~ factor(TARGTPOP)*factor(DISPOINC), data = DwaineData) # This is the full model
#summary(Full)

#anova (Reduced, Full)


```



```{r}
# Problem 5 from exercises

p5 <- read.table("http://users.stat.ufl.edu/~rrandles/sta4210/Rclassnotes/data/textdatasets/KutnerData/Chapter%20%206%20Data%20Sets/CH06PR05.txt", sep ="" , header = FALSE)


p5Data <- data.frame(p5)


colnames(p5Data) <- c("y", "x1", "x2")
(p5Data)
dim(p5Data)


Reduced <- lm(y ~ x1 + x2, data = p5Data) # This is our usual model
summary(Reduced)

Full <-  lm(y ~ factor(x1)*factor(x2), data = p5Data) # This is the full model
summary(Full)

```

Testing the hypothesis:

$H_0: \mbox{Reduced model is good}$ $\quad$ $H_a: \mbox{Full model is good}$

```{r}

anova (Reduced, Full)


```

