---
title: "Multiple Regression in $R$"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE, comment=NA, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We’ll use the `Credit` dataframe from the `ISLR` package to demonstrate multiple regression with:

A numerical outcome variable $y$, in this case credit card balance.

Two explanatory variables:

  1) A first numerical explanatory variable $x_1$. In this case, their credit limit.

  2) A second numerical explanatory variable $x_2$. In this case, their income (in thousands of dollars).

Note: This dataset is not based on actual individuals, it is a simulated dataset used for educational purposes.

Analogous to simple linear regression, where the regression function $E\{Y\} = \beta_0 + \beta_1X_1$ is
a line, regression function $E\{Y\} = \beta_0 + \beta_1X_1 +\beta_2X_2$ is a plane.

```{r comment=NA, warning=FALSE, message=FALSE}
library(ISLR)
library(plotly)

data(Credit)
head(Credit)

# draw 3D scatterplot
p <- plot_ly(data = Credit, z = ~Balance, x = ~Income, y = ~Limit, opacity = 0.6, color = Credit$Balance) %>%
  add_markers() 
p

#Fit a multiple linear regression model 
#Syntax: model_name <- lm(y ~ x1 + x2 + ... +xn, data = data_frame_name)

Balance_model <- lm(Balance ~ Limit + Income, data = Credit)
Balance_model

```

