---
title: "Final Capstone Project"
author: "Name:______________________"
date: "2024-08-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Certainly! I'll guide you through implementing the project in R using the Telco Customer Churn dataset from Kaggle. The project will cover data preprocessing, exploratory data analysis (EDA), building linear, logistic, and non-linear regression models, performing diagnostics, and comparing the models.

### Step 1: Setup and Data Loading

First, you'll need to download the dataset and load it into R.

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(caret)
library(car)
library(nnet)
library(MASS)
library(pROC)


# Set the file path (adjust the path if necessary)
file_path <- "WA_Fn-UseC_-Telco-Customer-Churn.csv"

# Load the dataset
telco_data <- read.csv(file_path)

# Preview the data
head(telco_data)
str(telco_data)
summary(telco_data)
```

### Step 2: Data Preprocessing

We need to clean the data, handle missing values, and convert categorical variables to factors.

```{r}
# Handle missing values
telco_data <- telco_data %>% na.omit()

# Convert categorical variables to factors
telco_data$Churn <- as.factor(telco_data$Churn)
telco_data$gender <- as.factor(telco_data$gender)
telco_data$Partner <- as.factor(telco_data$Partner)
telco_data$Dependents <- as.factor(telco_data$Dependents)
telco_data$PhoneService <- as.factor(telco_data$PhoneService)
telco_data$MultipleLines <- as.factor(telco_data$MultipleLines)
telco_data$InternetService <- as.factor(telco_data$InternetService)
telco_data$OnlineSecurity <- as.factor(telco_data$OnlineSecurity)
telco_data$OnlineBackup <- as.factor(telco_data$OnlineBackup)
telco_data$DeviceProtection <- as.factor(telco_data$DeviceProtection)
telco_data$TechSupport <- as.factor(telco_data$TechSupport)
telco_data$StreamingTV <- as.factor(telco_data$StreamingTV)
telco_data$StreamingMovies <- as.factor(telco_data$StreamingMovies)
telco_data$Contract <- as.factor(telco_data$Contract)
telco_data$PaperlessBilling <- as.factor(telco_data$PaperlessBilling)
telco_data$PaymentMethod <- as.factor(telco_data$PaymentMethod)
```

### Step 3: Exploratory Data Analysis (EDA)

We'll explore the data to understand relationships between features and the target variable (Churn).

```{r}
# Summary statistics of numeric features
summary(telco_data %>% select_if(is.numeric))

# Churn distribution
ggplot(telco_data, aes(x = Churn)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Distribution of Churn")

# Relationship between MonthlyCharges and Churn
ggplot(telco_data, aes(x = MonthlyCharges, fill = Churn)) +
  geom_histogram(binwidth = 5, alpha = 0.7, position = "identity") +
  labs(title = "Monthly Charges vs Churn")

# Relationship between Tenure and Churn
ggplot(telco_data, aes(x = tenure, fill = Churn)) +
  geom_histogram(binwidth = 5, alpha = 0.7, position = "identity") +
  labs(title = "Tenure vs Churn")
```

### Step 4: Linear Regression Model

We'll start by building a linear regression model to predict the probability of churn.

```{r}
# Linear regression model
linear_model <- lm(tenure ~ MonthlyCharges + TotalCharges + Contract, data = telco_data)

# Summary of the model
summary(linear_model)
```

### Step 5: Model Diagnostics

We will check for multicollinearity and plot residuals to diagnose any issues with the linear model.

```{r}
# Check multicollinearity
vif(linear_model)

# Residuals plot
plot(linear_model)
```

### Step 6: Logistic Regression

Since churn is a binary variable, we'll use logistic regression to predict churn.

```{r}
# Logistic regression model
logistic_model <- glm(Churn ~ MonthlyCharges + TotalCharges + Contract, data = telco_data, family = "binomial")

# Summary of the model
summary(logistic_model)

# Predicting probabilities
pred_probs <- predict(logistic_model, type = "response")

# ROC Curve and AUC
roc_curve <- roc(telco_data$Churn, pred_probs)
plot(roc_curve)
auc(roc_curve)
```

### Step 7: Non-Linear Regression

If non-linear relationships are suspected, a non-linear model can be implemented.

```{r}
# Example: Non-linear model using nnet package (Neural Network)
nonlinear_model <- nnet::nnet(Churn ~ MonthlyCharges + TotalCharges + Contract, data = telco_data, size = 5, linout = TRUE)

# Summary of the model
summary(nonlinear_model)
```

### Step 8: Model Comparison

We will compare the models using metrics like AIC, BIC, and misclassification error.



```{r}
# Compare AIC and BIC
AIC(linear_model, logistic_model)
BIC(linear_model, logistic_model)

# Misclassification Error for Logistic Model
pred_class <- ifelse(pred_probs > 0.5, "Yes", "No")
conf_matrix <- table(Predicted = pred_class, Actual = telco_data$Churn)
conf_matrix
misclass_error <- 1 - sum(diag(conf_matrix)) / sum(conf_matrix)
misclass_error
```

### Step 9: Final Report

You can compile the results into a report and create visualizations as needed.

```{r}
# Visualization of model comparison
ggplot(telco_data, aes(x = MonthlyCharges, y = as.numeric(Churn))) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  labs(title = "Logistic Regression Fit")

# Save the workspace
save.image(file = "churn_analysis.RData")
```

This code covers all aspects of the project. You can run each block in R, modify as needed, and interpret the outputs for your final report.